A More Complete Algebra for File Synchronization

### 2015-11-11 2015/paper.tex
Finished proof; added definitions for commands and documented derivations

\usepackage{alltt}
\usepackage{comment}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{setspace}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{afterpage}
\usepackage{centernot}

\usepackage{tikz}

\newcommand{\setv}{\mathcal{V}}
\newcommand{\setvx}[1]{\mathcal{V}_{#1}}
\newcommand{\setf}{\setvx{F}}
\newcommand{\setd}{\setvx{D}}
\newcommand{\setb}{\setvx{\empt}}
\newcommand{\setp}{\mathcal{P}}
\newcommand{\empt}{\bot}
\newcommand{\parent}{\mathtt{parent}}
\newcommand{\FS}{\Phi} % {\mathrm{FS}}
\newcommand{\GS}{\Xi} % another filesystem
\newcommand{\pp}{p^\uparrow} % parent path
\newcommand{\np}{p_{\centernot\leftrightarrow}} % path not related to p

@@

\theoremstyle{definition}
\newtheorem{mydef}{Definition}

\newtheorem{mylem}{Lemma}

@@

Every filesystem has a so-called \textbf{tree property}, which means that
if the filesystem is not empty at a path, and it is not a root path,
then it must have a directory at the path's parent.
This requires that a parent--child relationship is defined over $\setp$,
which we model using the function $\parent$, which returns the parent path
or $\empt$ if the path is a root path:
\[ \parent : \setp \rightarrow \setp \cup \{\empt\} \]

Moreover, in $\setv$ we distinguish between files ($\setf$) and directories ($\setd$), that is,
if $\setb = \{\empt\}$ then:
\[ \setv = \setb \cup \setf \cup \setd \]
The tree propety can then be defined as
\[ \forall \FS, p\in\setp : \FS(p) \neq \empt \wedge \parent(p) \neq \empt \Rightarrow \FS(\parent(p)) \in \setd \]

@@

In other words, the types will form a sequence like $D,D,\cdots,D,F,\empt,\empt,\cdots$.

@@

\section{Commands}

@@

Let us consider what kind of information may be encoded in the commands.
A usual set of commands, based on the most frequent tools implemented by filesystems,
might be $create(p,V)$, $edit(p,V)$ and $remove(p)$ where $p\in\setp$ and $V\in\setv$ (but $V\neq\empt$).
Clearly the commands need to contain information about the state they leave the filesystem
in at the path on which they operate, that is, they need to contain the ``end'' value, $V$.
Notice, however, that the commands above also encode some information about the filesystem
{\it before} the command is applied; namely, $create$ requires that there are no files
or directories at $p$, while $edit$ and $remove$ require the opposite.

@@

In line with the aim to encode in commands as much information as possible,
and in order to resolve this asymmetry, we propose a set of commands that encode
the type of the original state of $\FS(p)$ as well.
(Some real-life filesystem commands like $mkdir$ or $rmdir$ do this already.)
Please note that there is never any need to encode information about the
{\it exact} value of path $p$ in a command, merely its type ($D$, $F$ or $\empt$),
as the success or failure of subsequent commands only depend on the type of the value.

We therefore propose to have a command for each pair of types.
For want of a better system, we will simply name our commands by concatenating
two of $\fsregcommandchar{D}$, $\fsregcommandchar{F}$ and $\fsregcommandchar{\empt}$, 
where the left sign notes the type of value
in the filesystem before the command is applied, and the right sign notes the type
afterwards. For example $mkdir(p,D)$ is $\cbd(p,D)$ and $rmdir(p)$ is $\cdb(p)$.

@@

The commands can only succeed if the original value at $p$ has a type that matches
the type required by a command. If this is not the case, or if the resulting
filesystem no longer has the tree property, then we say that the command
\textbf{breaks} the filesystem. Broken filesystems are considered equal
(but not equal to any working filesystem), and we note them by $\FS=\empt$.

So that we could reason about sequences of commands that break every filesystem
in the algebra, we introduce the command $\cbrk$ that simply breaks every filesystem.
We note that a sequence of commands $S$ is applied to a filesystem
by prefixing the sequence to it:
\begin{itemize}
\item $S\FS$
\item $[\cbrk]\FS=\empt$
\item $[\cff(p,F)]\FS = \GS$, for which $\GS(p)=F$ holds
and $\GS(q)=\FS(q)$ for any path $q\neq p$.
\end{itemize}

@@

Pairs of commands in general will have the form
\[ [\cxy(p,V_Y); \czw(q,V_W)] \]
where $X,Y,Z,W\in\{\fsregcommandchar{D},\fsregcommandchar{F},\empt\}$, $p,q\in\setp$, and values are of the appropriate type: 
$V_Y\in\setvx{Y}$ and $V_W\in\setvx{W}$.
(Strictly speaking, commands where $Y=\empt$ or $W=\empt$ do not require a value,
but they can also be written in a $\fscommand{X}{\empt}(p, \empt)$ form.)

@@

We have two inference rules for the two judgements,
which make it possible to derive new statements from the axioms or other statements.
These are:

\emph{For any sequences \(S_1, S_2, S, S'\):}
\begin{itemize}
\item if \(S_1\equiv S_2\) then \(S;S_1;S'\equiv S;S_2;S'\)
\item if \(S_1\eqext S_2\) then \(S;S_1;S'\eqext S;S_2;S'\)
\item if \(S_1\equiv S_2\) then \(S_1\eqext S_2\).
\end{itemize}

Finally, we have a set of axioms or laws that describe the behaviour
of each command pair. We exclude commands that either break the
filesystem or leave it in the same state ($\cbb$) and for
simplicity's sake we merged axioms where the outcome is predictable
from the relationship between the paths in the commands.

\begin{description}
\item[Unrelated.] If the two paths in the pair are not directly related, then
the commands commute with an equivalence and they will not break every filesystem:
\begin{align*}
[\cxy(p,V_Y); \czw(\np,V_W)] &\equiv [\czw(\np,V_W); \cxy(p,V_Y)] \\
&\nequiv \cbrk
\end{align*}
%
\item[Same.] If the two paths are the same ($[\cxy(p,V_Y); \czw(p,V_W)]$), the
the commands cannot be reversed. Instead:
   \begin{description}
   \item[Break.]
   The pair will break every filesystem if $Z\ne Y$, that is, if the second
   command expects a type of value different from the one left by the first command.
   \item[Simplified by extension.]
   Otherwise, the pair will simplify to a single or no commands
   if both $X$ and $W$ are $\fsregcommandchar{\empt}$ or $\fsregcommandchar{D}$:
   \begin{gather*}
   % bf=fb Sp
            [\cbf(p, F); \cfb(p)] \eqext \emptyseq \\
   % bd=db Sp
            [\cbd(p, D); \cdb(p)] \eqext \emptyseq \\
   % db=bd Sp
            [\cdb(p); \cbd(p, D)] \eqext [\cdd(p, D)] \\
   % df=fd Sp
            [\cdf(p, F); \cfd(p, D)] \eqext [\cdd(p, D)]
   \end{gather*}
   \item[Simplified by equivalence.]
   In all other cases, the commands are equivalent to a single command
   replacing the input of the first command with the output of the second:
   \[ [\cxy(p, V_Y); \cyz(p, V_Z)] \equiv [\cxz(p, V_Z)] \]
   \end{description}
%
\item[Down.]
These are the pairs in which the first command is applied to the parent of the path
of the second command.
   \begin{description}
   \item[Construction.] There are pairs in this group to which no law applies.
   We call these ``construction'' pairs:
   \begin{gather*}
   % bd>bf
            [\cbd(\pp, D); \cbf(p, F)] \\
   % bd>bd
            [\cbd(\pp, D_1); \cbd(p, D_2)] \\
   % fd>bf
            [\cfd(\pp, D); \cbf(p, F)] \\
   % fd>bd
            [\cfd(\pp, D_1); \cbd(p, D_2)]
   \end{gather*}
   \item[Commute.] Otherwise, the commands in these pairs commute iff the first command is $\cdd$. For example:
   \begin{gather*}
   [\cdd(\pp, D); \cbf(p, F)] \equiv [\cbf(p, F); \cdd(\pp, D)]
   \end{gather*}
   \item[Break.] All other pairs break every filesystem, e.g.:
   \begin{gather*}
   [\cdf(\pp, F); \cbd(p, D)] \equiv \cbrk
   \end{gather*}
   \end{description}
%
\item[Up.]
The final group of pairs are the ones in which the path in the second
command is the parent of the one in the first.
These behave similarly to the previous group. 
   \begin{description}
   \item[Desctruction.] There are
   so-called ``destruction'' pairs to which no law applies:
   \begin{gather*}
   % fb<db ??
            [\cfb(p); \cdb(\pp)] \\
   % fb<df ??
            [\cfb(p); \cdf(\pp, F)] \\
   % db<db ??
            [\cdb(p); \cdb(\pp)] \\
   % db<df ??
            [\cdb(p); \cdf(\pp, F)]
   \end{gather*}
   \item[Commute.] Otherwise the commands commute iff the \emph{second} command is $\cdd$.
   \item[Break.] All other pairs break every filesystem.
   \end{description}
\end{description}

@@

In real-life applications, we often do not want to consider metadata stored in
directories (\cite{BZ}). In other words, the contents of directories are all equal,
which can be modelled by assuming that $|\setd|=1$.
Note that this creates a convenient symmetry where there are only a single value
for each type that can be repeated as we move from a path to its child ($D$ and $\empt$).

@@

{\bf Going forward we use the algebra derived from the $|\setd|=1$ case.}

@@

\begin{mydef}{Minimal set / sequence}
By a minimal sequence or set of commands we mean a sequence / set in which
there is at most one command on each path.
\end{mydef}

If $U$ is a minimal set of commands
that we know can be applied to at least one filesystem in some order without breaking it,
then we can split $U$ into a maximum number of disjunct subsets
(components) so that the paths of two commands from different subsets would never be directly related.
If we aim to order the commands in $U$, then we can focus on ordering the commands
inside the components, as all permutations of the components are allowed:
the commands in them will commute.

@@

In each component that is larger than one, the paths are directly related
and the commands form construction or destruction chains.
To order the commands,
select the command with the topmost path, so that there would be no command
on its parent path. If the command removes content at this path, then it
needs to follow all other commands in the component, and
the component is a destruction component.
Otherwise the command with the topmost component needs to precede other commands
in the component, and it is a construction component.
Repeat the process of splitting commands into subsets on the remaining commands
in the component. Recursively, this way we can determine all allowed orders of the commands.

\begin{mydef}{$\orderset{U}$}
We use $\orderset{U}$ to denote the set of sequences that contain all commands in $U$
and represent all allowed orders.
\end{mydef}

As a specific case, we can regard a sequence of commands $S$
as a set and determine its $\orderset{S}$.
At times by $\orderset{U}$ we mean an arbitrary sequence in the set.

\begin{mylem}\label{lemma:neighbour}
If $U$ contains commands on both $p$ and $\pp$, then
there is a $U^*\in\orderset{U}$ where they are next to each other.
\end{mylem}

\begin{mydef}
$\extset{S} := \{S^*|S\eqext S^*\}$
\end{mydef}

\begin{mylem}\label{lemma:minextset}
For any minimal sequence $S$ not equivalent to $\cbrk$, $\extset{S} = \{S^*|S\equiv S^*\}$.
\end{mylem}
\begin{proof}
This is because as $S$ has at most one command on each path, only equivalence axioms
can be applied to it.
\end{proof}

@@

@@

\begin{mydef}{Fully typed sequences}
We call a sequence of commands fully typed if whenever it breaks a filesystem
it is explained by a mismatch on the input types of the commands it contains.
TODO: define fully typed sequences in terms of the algebra?
\end{mydef}

To aid the construction of fully typed sequences, we reintroduce
two assertion commands that either break the filesystem
or leave it unchanged: $\cbb(p)$ and $\cdd(p)$.
As $|\setb|=|\setd|=1$, we do not note the exact output value.
These can be used to test whether $\FS(p)=\empt$ or $\FS(p)=D$.

@@

\begin{itemize}
\item $A\coworks B \Rightarrow A\nequiv\cbrk$
\item $A\eqext B \Rightarrow A\wrext B$
\item $A\coworks B\wrext C \Rightarrow A\coworks C$
\item $A\ncoworks B\wrnrw C \Rightarrow A\ncoworks C$
\item $A\ncoworks B \Rightarrow A;S\ncoworks B$
\end{itemize}

Axioms:

\begin{itemize}
\item $\cxy(p,V_Y)\ncoworks \czw(p,V_W)$ where $X\neq Z$
% \item Disjoint laws for construction:
%    \begin{gather*}
%    % bd>bf
%             [\cbd(\pp, D); \cbf(p, F)] \ncoworks [\cbf(p, F)]\\
%    % bd>bd
%             [\cbd(\pp, D_1); \cbd(p, D_2)] \ncoworks [\cbd(p, D_2)]\\
%    % fd>bf
%             [\cfd(\pp, D); \cbf(p, F)] \ncoworks [\cbf(p, F)]\\
%    % fd>bd
%             [\cfd(\pp, D_1); \cbd(p, D_2)] \ncoworks [\cbd(p, D_2)]
%    \end{gather*}
% \item Disjoint laws for destruction:
%    \begin{gather*}
%    % fb<db ??
%             [\cfb(p); \cdb(\pp)] \ncoworks [\cdb(\pp)]\\
%    % fb<df ??
%             [\cfb(p); \cdf(\pp, F)] \ncoworks [\cdf(\pp, F)]\\
%    % db<db ??
%             [\cdb(p); \cdb(\pp)] \ncoworks [\cdb(\pp)]\\
%    % db<df ??
%             [\cdb(p); \cdf(\pp, F)] \ncoworks [\cdf(\pp, F)]
%    \end{gather*}
\end{itemize}

@@

\begin{mylem}
If $A$ is fully typed and $S;A$ is minimal, then $S;A\wrext A$
TODO: Is this an axiom?
\end{mylem}

\begin{mylem}\label{lemma:ncowprefix}
$A\ncoworks B \Rightarrow S;A\ncoworks B$ if $A$ is fully typed and $S;A$ is minimal.
\end{mylem}
\begin{proof}
This follows from above.
\end{proof}

@@

\begin{mylem}
Let $A$ and $B$ be two minimal sequences so that $A\coworks B$.
Then their intersection can also be applied to any filesystem
either can be applied to without breaking it.
\end{mylem}

@@

In terms of the algebra, this latter argument could be rephrased as follows.
Split $A$ into $[A_0;\czw(q, V_W);\cxy(p, V_Y);A_1]$.
Add command $\cqq(q)$ before $\cxy(p, V_Y)$ to $B$ according to the construction or destruction typing laws.
This yields $B^*\equiv B$, which we split into $[B_0;\cqq(q);\cxy(p, V_Y);B_1]$.
We know $[\czw(q, V_W);\cxy(p, V_Y)]$ is fully typed (TODO: why?)
and $A$ is minimal, so $[A_0;\czw(q, V_W);\cxy(p, V_Y)]$ is minimal.
We also know $[\cqq(q);\cxy(p, V_Y)]$ is fully typed and $B^*$ is minimal, so $[B_0;\cqq(q);\cxy(p, V_Y)]$ is minimal.
But we know $[\czw(q, V_W);\cxy(p, V_Y)]\ncoworks [\cqq(q);\cxy(p, V_Y)]$
as comparing the construction and destruction laws to their typing counterparts we know that $Z\neq Q$,
and so from lemma \ref{lemma:ncowprefix} and $A\ncoworks B \Rightarrow A;S\ncoworks B$
we know that $A\ncoworks B^*\equiv B$ and so $A\ncoworks B$ which is a contradiction.
(TODO: Is there a point in proving this this way?)

@@

\end{proof}

>>>>

\usepackage{alltt}
\usepackage{comment}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{setspace}
% \usepackage{fullpage}
\usepackage{graphicx}
% \usepackage{afterpage}

\usepackage{centernot}
\usepackage{tikz}

\newcommand{\setv}{\mathcal{V}}
\newcommand{\setvx}[1]{\mathcal{V}_{#1}}
\newcommand{\setf}{\setvx{F}}
\newcommand{\setd}{\setvx{D}}
\newcommand{\setb}{\setvx{\empt}}
\newcommand{\setp}{\mathcal{P}}
\newcommand{\empt}{\bot}
\newcommand{\parent}{\mathtt{parent}}
\newcommand{\toppath}{\mathtt{None}}
\newcommand{\fsbroken}{\mathtt{Broken}}
\newcommand{\FS}{\Phi} % {\mathrm{FS}}
\newcommand{\GS}{\Xi} % another filesystem
\newcommand{\pp}{p^\uparrow} % parent path
\newcommand{\np}{p_{\centernot\leftrightarrow}} % path not related to p

@@

\theoremstyle{definition}
\newtheorem{mydef}{Definition}
\newtheorem{myax}{Axiom}
\newtheorem{mylem}{Lemma}

@@

The paths of a filesystem form a disjoint union of rooted directed trees.
\begin{mydef}
The function $\parent(p)$ returns the parent path of $p$, or
returns $\toppath$ if $p$ is the root of a tree.
We also write $\pp$ for $\parent(p)$.
\end{mydef}

Every filesystem has a so-called \textbf{tree property}, which means that
if the filesystem is not empty at a path, and the path has a parent,
then there must be a directory at the parent path.

To model this, in $\setv$ we distinguish between files ($\setf$) and directories ($\setd$), that is,
if $\setb = \{\empt\}$ then:
\[ \setv = \setb \cup \setf \cup \setd \]
The tree propety can then be defined as
\[ \forall \FS, p\in\setp : \FS(p) \neq \empt \wedge \parent(p) \neq \toppath \Rightarrow \FS(\parent(p)) \in \setd \]

@@

@@

\section{Introducing commands}

@@

Let us consider what kind of information may be encoded in the commands.
A usual set of commands, based on the most frequent tools implemented by filesystems,
might be $create(p,V)$, $edit(p,V)$ and $remove(p)$ where $p\in\setp$ and $V\in\setv$ (but $V\neq\empt$).
Clearly the commands need to contain information about the state they leave the filesystem
in at the path on which they operate, that is, they need to contain their output value, $V$.
Notice, however, that the commands above also encode some information about the filesystem
\emph{before} the command is applied; namely, $create$ requires that there are no files
or directories at $p$, while $edit$ and $remove$ require the opposite.

@@

In line with the aim to encode as much information as possible in the commands,
and in order to resolve this asymmetry, we propose a set of commands that encode
the type of the original state of $\FS(p)$ as well.
(Some real-life filesystem commands like $rmdir$ do this already.)
Please note that there is never any need to encode information about the
\emph{exact} value of path $p$ in a command, merely its type ($D$, $F$ or $\empt$),
as the success or failure of filesystem commands only depend on the type of the value.

We therefore propose to have a command for each pair of types.
For want of a better system, we will name our commands by concatenating
two of $\fsregcommandchar{D}$, $\fsregcommandchar{F}$ and $\fsregcommandchar{\empt}$, 
where the left sign notes the type of value
in the filesystem before the command is applied, and the right sign notes the type
afterwards. For example $mkdir(p,D)$ is $\cbd(p,D)$ and $rmdir(p)$ is $\cdb(p)$.
Please find their exact definitions below.

@@

The commands can only succeed if the original value at $p$ has a type that matches
the type required by a command. If this is not the case, or if the resulting
filesystem no longer has the tree property, then we say that the command
\textbf{breaks} the filesystem. Broken filesystems are considered equal
(but not equal to any working filesystem), and we note them by $\FS=\fsbroken$.

So that we could reason about sequences of commands that break every filesystem
in the algebra, we introduce the command $\cbrk$ that simply breaks every filesystem.

We note that a command or a sequence of commands is applied to a filsystem
by prefixing the command or sequence to it, for example: $\cbrk\FS$, $\cbd(p,D)\FS$, 
or $S\FS$ if $S$ is a sequence of commands.

The exact behaviour of our commands is as follows:
\begin{itemize}
\item $\cbrk\FS = \fsbroken$
\item
All commands have the generic form $\cxy(p,V_Y)$ where
$X$ and $Y$ are types ($\fsregcommandchar{D}$, $\fsregcommandchar{F}$ or $\fsregcommandchar{\empt}$),
$p\in\setp$ and $V_Y$ is a value of the appropriate type from $\setvx{Y}$.
At times we omit $V_Y$ if there is only one suitable value.

Let us introduce the following notation for a filesystem derived from $\FS$:
\[ \FS\{p\mapsto V\}(q) :=
   \begin{cases}
   V &\mbox{if~} q=p\\
   \FS(q) &\mbox{otherwise.}
   \end{cases}
\]

Then
\[ \cxy(p,V_Y)\FS = 
   \begin{cases}
   \fsbroken &\mbox{if~} \FS=\fsbroken\\
   \fsbroken &\mbox{if~} \FS(p)\not\in\setvx{X}\\
   \fsbroken &\mbox{if~} \FS\{p\mapsto V_Y\} \mbox{~violates the tree property}\\
   \FS\{p\mapsto V_Y\} &\mbox{otherwise.}
   \end{cases}
\]
\end{itemize}

@@

Pairs of commands in general will have the form
\[ [\cxy(p,V_Y); \czw(q,V_W)] \]
where $X,Y,Z,W\in\{\fsregcommandchar{D},\fsregcommandchar{F},\empt\}$, $p,q\in\setp$, 
and values are of the appropriate type: 
$V_Y\in\setvx{Y}$ and $V_W\in\setvx{W}$.

@@

We have inference rules for the two judgements,
which make it possible to derive new statements from the axioms or other statements.
These are:

\emph{For any sequences \(S_1, S_2, S, S'\):}
\begin{itemize}
\item \(S_1\equiv S_2 \Rightarrow S;S_1;S'\equiv S;S_2;S'\)
\item \(S_1\eqext S_2 \Rightarrow S;S_1;S'\eqext S;S_2;S'\)
\item \(S_1\equiv S_2 \Rightarrow S_1\eqext S_2\).
\end{itemize}

Finally, we have a set of axioms or laws that describe the behaviour
of each command pair. We exclude commands that either break the
filesystem or leave it in the same state ($\cbb$) and for
simplicity's sake instead of listing axioms for all pairs of commands
we describe the behaviour of groups of pairs
based on the relationship between the paths in the commands.

\begin{description}
\item[Unrelated.] If the two paths in the pair are not directly related, then
the commands commute with an equivalence and they will not break every filesystem:
\begin{align*}
[\cxy(p,V_Y); \czw(\np,V_W)] &\equiv [\czw(\np,V_W); \cxy(p,V_Y)] \\
&\nequiv \cbrk
\end{align*}
%
\item[Same.] If the two paths are the same ($[\cxy(p,V_Y); \czw(p,V_W)]$), the
the commands cannot be reversed. Instead:
   \begin{description}
   \item[Break.]
   The pair will break every filesystem if $Z\ne Y$, that is, if the second
   command expects a type of value different from the one left by the first command.
   \item[Simplified by extension.]
   Otherwise, the pair will simplify to a single or no commands
   if both $X$ and $W$ are $\fsregcommandchar{\empt}$ or $\fsregcommandchar{D}$:
   \begin{gather*}
   % bf=fb Sp
            [\cbf(p, F); \cfb(p)] \eqext \emptyseq \\
   % bd=db Sp
            [\cbd(p, D); \cdb(p)] \eqext \emptyseq \\
   % db=bd Sp
            [\cdb(p); \cbd(p, D)] \eqext [\cdd(p, D)] \\
   % df=fd Sp
            [\cdf(p, F); \cfd(p, D)] \eqext [\cdd(p, D)]
   \end{gather*}
   \item[Simplified by equivalence.]
   In all other cases, the commands are equivalent to a single command
   replacing the input of the first command with the output of the second:
   \[ [\cxy(p, V_Y); \cyz(p, V_Z)] \equiv [\cxz(p, V_Z)] \]
   \end{description}
%
\item[Down.]
These are the pairs in which the first command is applied to the parent of the path
of the second command.
   \begin{description}
   \item[Construction.] There are pairs in this group to which no law applies:
   the commands do not commute or break every filesystem, and cannot be simplified.
   We call these ``construction'' pairs:
   \begin{gather*}
   % bd>bf
            [\cbd(\pp, D); \cbf(p, F)] \\
   % bd>bd
            [\cbd(\pp, D_1); \cbd(p, D_2)] \\
   % fd>bf
            [\cfd(\pp, D); \cbf(p, F)] \\
   % fd>bd
            [\cfd(\pp, D_1); \cbd(p, D_2)]
   \end{gather*}
   \item[Commute.] Otherwise, the commands in these pairs commute iff the first command is $\cdd$. For example:
   \begin{gather*}
   [\cdd(\pp, D); \cbf(p, F)] \equiv [\cbf(p, F); \cdd(\pp, D)]
   \end{gather*}
   \item[Break.] All other pairs break every filesystem, e.g.:
   \begin{gather*}
   [\cdf(\pp, F); \cbd(p, D)] \equiv \cbrk
   \end{gather*}
   \end{description}
%
\item[Up.]
The final group of pairs are the ones in which the path in the second
command is the parent of the one in the first.
These behave similarly to the previous group. 
   \begin{description}
   \item[Desctruction.] There are
   so-called ``destruction'' pairs to which no law applies:
   \begin{gather*}
   % fb<db ??
            [\cfb(p); \cdb(\pp)] \\
   % fb<df ??
            [\cfb(p); \cdf(\pp, F)] \\
   % db<db ??
            [\cdb(p); \cdb(\pp)] \\
   % db<df ??
            [\cdb(p); \cdf(\pp, F)]
   \end{gather*}
   \item[Commute.] Otherwise the commands commute iff the \emph{second} command is $\cdd$.
   \item[Break.] All other pairs break every filesystem.
   \end{description}
\end{description}

@@

In real-life applications, we often do not want to consider metadata stored in
directories (\cite{BZ}). In other words, the contents of directories are all equal,
which can be modelled by assuming that $|\setd|=1$.
Note that this creates a convenient symmetry where there is only a single value
for each type that can be repeated as we move from a path to its child ($D$ and $\empt$).

@@

% TODO
{\bf Going forward we use the algebra derived from the $|\setd|=1$ case.}

@@

\begin{mydef}{Minimal set / sequence}
By a minimal set or sequence of commands we mean a set / sequence in which
there is at most one command on every path.
\end{mydef}

If $U$ is a minimal set of commands
that we know can be applied to at least one filesystem in some order without breaking it,
then we can split $U$ into a maximum number of disjunct subsets
(components) so that the paths of two commands from two different subsets are never directly related.
If we aim to order the commands in $U$, then we can allow any permutation of 
the components as the commands in them will commute freely.

@@

In each component that is larger than one, the paths 
form a rooted ordered tree via the parent--child relationships,
and the pairs connected by edges will be like pairs in the
construction or destruction laws.
To order the commands in a component,
select the command with the topmost path, so that there would be no command
on its parent path (if there is one). 
If the command removes content at this path (it is of the form $\fsregcommand{X}{\empt}$), then it
needs to follow all other commands in the component, and
the component is made of destruction pairs.
Otherwise the command with the topmost component needs to precede other commands
in the component, and it is made of construction pairs.
Move the command with the topmost path in place, and
repeat the process of splitting commands into subsets on the remaining commands
in the component. Recursively, this way we can determine all allowed orders of the commands.

\begin{mydef}{$\orderset{U}$}
We use $\orderset{U}$ to denote the set of sequences that contain all commands in $U$
and represent all orders generated by this algorithm.
At times by $\orderset{U}$ we mean an arbitrary sequence in the set.
\end{mydef}

As a specific case, we can regard a sequence of commands $S$
as a set and determine its $\orderset{S}$.

\begin{mylem}\label{lemma:neighbour}
If $U$ contains commands on both $p$ and $\pp$, then
there is a $U'\in\orderset{U}$ where they are next to each other.
\end{mylem}

\begin{mydef}
$\extset{S} := \{S'|S\eqext S'\}$
\end{mydef}

\begin{mylem}\label{lemma:minextset}
For any minimal sequence $S$ not equivalent to $\cbrk$, $\extset{S} = \{S'|S\equiv S'\}$.
\end{mylem}
\begin{proof}
This is because as $S$ has at most one command on each path, only equivalence axioms
can be applied to it.
\end{proof}

@@

% TODO Can these be derived from the algebra?

@@

\begin{mydef}{Fully typed sequences}
We call a sequence of commands fully typed if whenever it breaks a filesystem
it is explained by a mismatch on the input types of the commands it contains.
% TODO
TODO: define fully typed sequences in terms of the algebra?
\end{mydef}

To aid the construction of fully typed sequences, we reintroduce
two assertion commands that either break the filesystem
or leave it unchanged: $\cbb(p)$ and $\cdd(p)$.
(As $|\setb|=|\setd|=1$, we do not note their exact output value.)
They can be used to test whether $\FS(p)=\empt$ or $\FS(p)=D$.

@@

\begin{myax}$A\coworks B \Rightarrow A\nequiv\cbrk$\end{myax}
\begin{myax}$A\eqext B \Rightarrow A\wrext B$\end{myax}
\begin{myax}\label{axiom:wrextpostfix}$[A;S] \wrext A$\end{myax}
\begin{myax}$A\coworks B\wrext C \Rightarrow A\coworks C$\end{myax}
\begin{myax}\label{axiom:ncowwrnrw}$A\ncoworks B\wrnrw C \Rightarrow A\ncoworks C$\end{myax}

\begin{mylem}\label{lemma:ncowpostfix}$A\ncoworks B \Rightarrow [A;S]\ncoworks B$\end{mylem}
\begin{proof}From axioms \ref{axiom:wrextpostfix} and \ref{axiom:ncowwrnrw}\end{proof}

\begin{myax}$\cxy(p,V_Y)\ncoworks \czw(p,V_W)$ if $X\neq Z$\end{myax}
% \item Disjoint laws for construction:
%    \begin{gather*}
%    % bd>bf
%             [\cbd(\pp, D); \cbf(p, F)] \ncoworks [\cbf(p, F)]\\
%    % bd>bd
%             [\cbd(\pp, D_1); \cbd(p, D_2)] \ncoworks [\cbd(p, D_2)]\\
%    % fd>bf
%             [\cfd(\pp, D); \cbf(p, F)] \ncoworks [\cbf(p, F)]\\
%    % fd>bd
%             [\cfd(\pp, D_1); \cbd(p, D_2)] \ncoworks [\cbd(p, D_2)]
%    \end{gather*}
% \item Disjoint laws for destruction:
%    \begin{gather*}
%    % fb<db ??
%             [\cfb(p); \cdb(\pp)] \ncoworks [\cdb(\pp)]\\
%    % fb<df ??
%             [\cfb(p); \cdf(\pp, F)] \ncoworks [\cdf(\pp, F)]\\
%    % db<db ??
%             [\cdb(p); \cdb(\pp)] \ncoworks [\cdb(\pp)]\\
%    % db<df ??
%             [\cdb(p); \cdf(\pp, F)] \ncoworks [\cdf(\pp, F)]
%    \end{gather*}

@@

\begin{mylem}\label{lemma:wrextprefix}
$[S;A]\wrext A$
if $A$ is fully typed and $[S;A]$ is minimal.
TODO: Is this an axiom?
\end{mylem}

\begin{mylem}\label{lemma:ncowprefix}
$A\ncoworks B \Rightarrow [S;A]\ncoworks B$ 
if $A$ is fully typed and $[S;A]$ is minimal.
\end{mylem}
\begin{proof}
This follows from lemma \ref{lemma:wrextprefix} and axiom \ref{axiom:ncowwrnrw}.
\end{proof}

@@

\begin{mylem}
Let $A$ and $B$ be two minimal sequences so that $A\coworks B$.
Then their intersection ($\orderset(A\cap B)$) can also be applied to any filesystem
either can be applied to without breaking it.
\end{mylem}

@@

In terms of the algebra, this latter argument could be rephrased as follows.
Split $A$ into $[A_0;\czw(q, V_W);\cxy(p, V_Y);A_1]$.
Add command $\cqq(q)$ before $\cxy(p, V_Y)$ to $B$ according to the construction or destruction typing laws.
This yields $B^*\equiv B$, which we split into $[B_0;\cqq(q);\cxy(p, V_Y);B_1]$.
We know $[\czw(q, V_W);\cxy(p, V_Y)]$ is fully typed (TODO: why?)
and $A$ is minimal, so $[A_0;\czw(q, V_W);\cxy(p, V_Y)]$ is minimal.
We also know $[\cqq(q);\cxy(p, V_Y)]$ is fully typed and $B^*$ is minimal, so $[B_0;\cqq(q);\cxy(p, V_Y)]$ is minimal.
But we know $[\czw(q, V_W);\cxy(p, V_Y)]\ncoworks [\cqq(q);\cxy(p, V_Y)]$
as comparing the construction and destruction laws to their typing counterparts we know that $Z\neq Q$,
and so from lemma \ref{lemma:ncowprefix} and lemma \ref{lemma:ncowpostfix}
we know that $A\ncoworks B^*\equiv B$ and so $A\ncoworks B$ which is a contradiction.
(TODO: Is there a point in proving this this way?)

@@

We now know that $A$ can be rearranged in a way that commands in $A\cap B$
are at the beginning, and therefore $A\wrext \orderset(A\cap B)$ and by symmetry $B\wrext \orderset(A\cap B)$.
\end{proof}
